---
title: "Forcasting Bitcoin Price Using ARIMA Model"
author: "Longyin Poon"
date: "4/10/2018"
output: pdf_document
---

```{r, echo=FALSE,results='hide',warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

#Training dataset includes 230 weeks from 4/8/2013 to 8/21/2017

#Testing dataset consists 30 weeks of data from 8/28/2017 to 3/26/2018

```{r, warning=FALSE,results="hide",echo=FALSE,message=FALSE}
#Load packages
library(jsonlite)
library(dplyr)
library(tidyr)
library(lubridate)
library(readr)
library(stringr)
library(ggplot2)
library(astsa)
library(tseries)
library(forecast)
```
```{r,message=FALSE}
#Load dataset
bitcoin_data <- read_csv("bitcoin_data.csv")
training <- bitcoin_data[1:230,]
testing <- bitcoin_data[231:260,]
```

#Time Series Analysis
```{r,echo=FALSE}
bitcoin_ts <- ts(training$bitcoin_price)

```

###Statitonary Test
```{r}
#Statitonary Test
tsplot(x=training$week,y=training$bitcoin_price, ylab = "Bitcoin Price", xlab = "Number of Weeks",main = "Weekly Bitcoin Price from 4/8/2013 to 8/21/2017", col="steelblue", lwd=3)
adf.test(bitcoin_ts, alternative = "stationary")
#Fail to reject the series is non-stationary with high p-value
```

###Log transformation
```{r}
training$log_price <- log(training$bitcoin_price)
tsplot(x=training$week,y=training$log_price, ylab = "Logged Bitcoin Price", xlab = "Number of Weeks",main = "Weekly Logged Bitcoin Price from 4/8/2013 to 8/21/2017", col="steelblue", lwd=3)

#The augmented Dickey-Fuller test
adf.test(training$log_price, alternative = "stationary")
#Fail to reject the series is non-stationary but p-value is lower
```

###ACF nd PACF for logged Bitcoin Price
```{r}
Acf(training$log_price, main="ACF for Logged Bitcoin Price") 
#Significant autocorrelations with many lags caused by carry over from earlier lags
Pacf(training$log_price, main="PACF for Logged Bitcoin Price" ) 
#A spike at 1 in PACF meaning lag 1 alone explained most of the information

```

###First order differencing on logged Bitcoin Price

```{r}
diff_log_price <- diff(training$log_price)
plot(diff_log_price, type='l',xlab="Number of Weeks", ylab="Differenced logged Bitcoin Price", main = "First Order Differencing on logged Bitcoin Price")
adf.test(diff_log_price, alternative = "stationary")
#Reject H0 with low p-value. The series is stationary
```

###ACF and PACF for Differenced log Bitcoin Price
```{r}
#ACF and PACF for differenced series to determine p and q for ARIMA model
Acf(diff_log_price, main='ACF for Differenced log Price')
#Significant spike at lag 1 and lag 2. Consider MA(2) model
Pacf(diff_log_price, main='PACF for Differenced log Price')
#Significant spike at lag 1 and close to significant for lag 2. Consider AR(1) or AR(2) model
```

###Scatterplot of logged Bitcoin Price against its lags
```{r}
plot(x=lag(training$log_price[2:230]), y=training$log_price[2:230], ylab = "Logged Bitcoin Price", xlab = "Lag 1 Logged Bitcoin Price", pch=19, cex=0.7, col = "steelblue", main = "Scatterplot of Logged Bitcoin Price VS Lag 1 Logged Bitcoin Price")

#Linear pattern suggesting the first order autoregressive model is appropriate.
```

#ARIMA Model
```{r,echo=FALSE}
fit<-auto.arima(training$log_price, seasonal=FALSE)
fit
#Auto arima function suggests (2,1,0) model
#AIC=-398.62   AICc=-398.46   BIC=-384.39
tsdisplay(residuals(fit), lag.max=30, main='(2,1,0) Model Residuals')
#No significant autocorrelation presents. Residuals are somewhat normally distributed.

fit_ma1 <- Arima(training$log_price,order = c(1,1,2))
fit_ma1
#AIC=-394.24   AICc=-394.08   BIC=-380.01
#ARIMA (1,1,2) has higher AIC , AICc and BIC

tsdisplay(residuals(fit_ma1), lag.max=30, main='(1,1,2) Model Residuals')

#No significant autocorrelation presents. Residuals are somewhat normally distributed.
```

###Check Residuals for model validity
```{R}
res <- fit$residuals
Acf(res, main="ACF for Residuals")
Pacf(res, main="PACF for Residuals")
#Based on ACF and PACF, residuals are randomly distributed.

Box.test(res,type="Ljung",lag=1)
#P-value is high so we fail to reject H0: The residuals are independent. Box-Ljung test confirms our results.

```

###Antilog function to reverse log transformation
```{r}
#Anti log function
antilog<-function(lx,base) { 
  lbx<-lx/log(exp(1),base=base) 
  result<-exp(lbx) 
  result 
} 
```

###Forecast
```{r}
fcast <- forecast(fit, h=30)
plot(fcast)

#Take the exponent
fcast$mean <- antilog(fcast$mean)
fcast$lower <- antilog(fcast$lower)
fcast$upper <- antilog(fcast$upper)
fcast$x <- antilog(fcast$x)

plot(fcast)
```

###Compare prediction and actual bitcoin price for 30 weeks
```{r}
actual <- testing$bitcoin_price
pred <- fcast$mean[1:30]
error <- abs(actual-pred)/actual
prediction_table <- cbind(actual,pred,error)
colnames(prediction_table) <- c("Actual","Prediction","%Change")
prediction_table <- as.data.frame(prediction_table)
prediction_table$Week <- seq(1:30)
prediction_table
mean(prediction_table$`%Change`)

ggplot() + 
  geom_line(aes(x=Week,y=Prediction,color="steelblue"), prediction_table)+ 
  geom_line(aes(x=Week, y=Actual,color="red"), prediction_table) +
  ylim(0,20000) +
  ylab("Bitcoin Price") +
  scale_color_manual(labels = c("Actual", "Prediction"), values = c("steelblue", "red")) + theme_classic() +
  ggtitle("Actual VS Predicted values")
  
```

#Add Google Trend As a Predictor
```{r}
tsdata <- ts(training)
fit_trend <- auto.arima(training$log_price, xreg = training$google_trend)
fit_trend #AIC=-405.75   AICc=-405.52   BIC=-387.97
#Lower AIC, AICc and BIC compare with model without google trend

tsdisplay(residuals(fit_trend), lag.max=30, main='(1,1,1) Model Residuals')
```

###Check Residuals for model validity
```{R}
res1 <- fit_trend$residuals
Acf(res1, main="ACF for Residuals")
Pacf(res1, main="PACF for Residuals")
#Based on ACF and PACF, residuals are randomly distributed.

Box.test(res1,type="Ljung",lag=1)
#P-value is high so we fail to reject H0: The residuals are independent. Box-Ljung test confirms our results.

```

###Google Trend ARIMA Model Forcast
```{r}
#Forecast
tsdata_trend <- ts(testing$google_trend)

fc.c2 <- forecast(tsdata_trend, h = 30)

newxreg <- as.matrix(fc.c2$mean)

fcast_trend <- forecast(fit_trend, xreg = newxreg)

plot(fcast_trend)

#Take the exponent
fcast_trend$mean <- antilog(fcast_trend$mean)
fcast_trend$lower <- antilog(fcast_trend$lower)
fcast_trend$upper <- antilog(fcast_trend$upper)
fcast_trend$x <- antilog(fcast_trend$x)


plot(fcast_trend)
```


###Compare prediction and actual bitcoin price for 30 weeks
```{r}
actual <- testing$bitcoin_price
pred1 <- fcast_trend$mean[1:30]
error1 <- abs(actual-pred1)/actual
prediction_table1 <- cbind(actual,pred1,error1)
colnames(prediction_table1) <- c("Actual","Prediction","%Change")
prediction_table1 <- as.data.frame(prediction_table1)
prediction_table1$Week <- seq(1:30)
prediction_table1
mean(prediction_table1$`%Change`)

ggplot() + 
  geom_line(aes(x=Week,y=Prediction,color="steelblue"), prediction_table1)+ 
  geom_line(aes(x=Week, y=Actual,color="red"), prediction_table1) +
  ylim(0,20000) +
  ylab("Bitcoin Price") +
  scale_color_manual(labels = c("Actual", "Prediction"), values = c("steelblue", "red")) + theme_classic() +
  ggtitle("Actual VS Predicted values")

```


#Accuracy of all ARIMA Model
```{r}
#Accuracy
accuracy(prediction_table$Actual,prediction_table$Prediction)
accuracy(prediction_table1$Actual,prediction_table1$Prediction)
```



